---
title: "Assignment_Three Q2"
author: "Author - Eric Opoku"
date: "December 8, 2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
fontsize: 12pt
geometry: margin=1in
keep_tex: yes
---


```{r setup, include=FALSE, comment=NA}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.width = 8,
	message = FALSE,
	warning = FALSE,
	collapse = TRUE
)
```


```{r header, echo=TRUE, message=FALSE, warning=FALSE, comment=NA}


name<- Sys.info()
name[7]
```



```{r message=FALSE, warning=FALSE, comment=NA}

library(rlang)
library(dplyr)
library(kableExtra)
library(ggdag)      # For plotting DAGs
library(dagitty)    # For working with DAG logic
library(modelsummary) # For making regression tables
library(AER) # this package has lots of applied metrics packages
library(foreign) 
# Helpful for reading in data from Stata or other code languages 
library(lubridate) # For figures
library(stargazer) # For tables
library(data.table)  #data manipulation and wrangling
library(lme4)
library(psych)
library(readxl) # Read in data
library(expss) #value labelling from spss style
library(readstata13)
library(marginaleffects) # To calculate marginal effects
library(knitr) # Alternative table package
library(ggplot2)





######
### 0. Load the packages we will need for this file ####
library(tidyverse) # load the installed package for each new session of R
library(broom)
library(causaldata) # Useful toy data sets
library(here) # Helpful in working with directories and projects
library(zoo) # Helpful packages for organizing dates
library(tidysynth) # For synthetic controls
library(gsynth) # For synthetic controls
library(gghighlight) # For figures 
library(binsreg) # For binscatters
library(nprobust) # Local linear regression

library(fixest)

set.seed(032620) # random number generators; same numbers across machines
```


#QUESTION ONE


```{r message=FALSE, warning=FALSE, comment=NA}
mydata<- read_excel("C:/Users/erico/Desktop/ACADEMICS/MY UofT COURSES/YEAR TWO/SEMESTER ONE/HAD5744H Applied Health Econometrics I/ASSIGNMENTS/ASSIGNMENT 3/a3_p1_sourdough_trends.xlsx") 

#View(mydata)


#regdata$the_date<-mydata$date  # I will use this data later


  
table(mydata$date)
table(mydata$hits)
table(mydata$keyword)
table(mydata$time)
table(mydata$gprop)
table(mydata$category)
table(mydata$geo)

mydata<-mydata%>%mutate(date=as.Date(date, format="YYYY-MM-DD"))
#View(mydata)

```


#QUESTION ONE A:
(i)shape
(ii)concerned about any control group? Yes, Soup and sandwich seems to have been affected by the pandemic but cereal is horizontal and seems like a good control variable

```{r Line graph, message=FALSE, warning=FALSE, comment=NA}


# Group to month level
#mydata <- mydata %>% group_by(date, keyword) %>% summarize(hits = sum(hits,na.rm=T))


#Show hits over time for each Keyword
ggplot(mydata,aes(x=date,y=hits,group=keyword, col=keyword)) + geom_line() + 
  theme_classic() + labs(x="Time",y="Monthly Hits")+ 
  geom_vline(xintercept = as.Date('2020-03-15'),color='black')




```


## QUESTION ONE B:


```{r Prior trends, message=FALSE, warning=FALSE, comment=NA}
table(mydata$keyword)

#View(mydata)
# Definition of key variables
mydata <- mydata %>% mutate(treated = ifelse(keyword == "sourdough", 1, 0), 
                              interaction = ifelse(keyword == "sourdough",date,0))

# Statistical tests: are the two groups' trends different?
pretrend_t <- lm(hits ~ date + interaction, data=mydata)

#using better control variable and dropping others
# Regression data -- just cereal and sourdough 
mydata <- mydata %>% filter(keyword %in% c("sourdough","cereal"))

pretrend_test <- lm(hits ~ date + interaction, data=mydata)
msummary(list(pretrend_t,pretrend_test),
         vcov=c(rep("robust",2)),
         stars=c('*' = .1, '**' = .05, '***' = .01)) # Interpret each coefficient here

```
## QUESTION ONE C:


```{r Relative month, message=FALSE, warning=FALSE, comment=NA}
# Definition of key variables

mydata$date <- as.yearmon(mydata$date)


mydata <- mydata %>% mutate(the_keyword = ifelse(keyword == "sourdough",1,0))
mydata <- mydata %>% mutate(After = ifelse(date >= "Mar 2020",1,0))
#mydata <- mydata %>% mutate(inter = keyword * post)
                            
table(mydata$hits)
table(mydata$After)
table(mydata$the_keyword)


# transform the data #Run pre trends again?
mydata <- mydata %>% mutate(logy = log(hits))


# First, construct relative time variable
mydata <- mydata %>% mutate(relative_month = round((as.numeric(date) - 2020.211)*12)) 

#View(mydata)




```

#QUESTION ONE D

```{r TWFE, message=FALSE, warning=FALSE, comment=NA}

#difference-in-difference estimate of the effect of lockdown on sourdough popularity.


dte <- feols(logy ~ i(relative_month, the_keyword, ref = -1) | 
                date + keyword, data = mydata)

# And use coefplot() for a graph of effects
iplot(dte,ref.line=-0.2, 
         ref.line.par=list(col="red",lty=1,lwd=2))


```




 


#QUESTION TWO

#A

```{r Synth control, message=FALSE, warning=FALSE, comment=NA}

mydata2<- read_excel("C:/Users/erico/Desktop/ACADEMICS/MY UofT COURSES/YEAR TWO/SEMESTER ONE/HAD5744H Applied Health Econometrics I/ASSIGNMENTS/ASSIGNMENT 3/a3_p2_organ_donations.xlsx") 

#View(mydata2)


```

```{r message=FALSE, warning=FALSE, comment=NA}

summary(mydata2$LivingDonors)

summary(mydata2$AllDonors)


summary(mydata2$WaitinglistAdditions)


```

```{r rates of living organ donations, message=FALSE, warning=FALSE, comment=NA}
#population-adjusted
mydata2$LivingDonors<-mydata2$LivingDonors/mydata2$Population*1000000
mydata2$AllDonors<-mydata2$AllDonors/mydata2$Population*1000000
mydata2$WaitinglistAdditions<-mydata2$WaitinglistAdditions/mydata2$Population*1000000

mydata2$GDP_percapita<-mydata2$GDP/mydata2$Population*1000000

mydata2$GDP_percapita_mean<-mean(mydata2$GDP_percapita)

#install.packages("table1")
library(table1)
table1(~GDP_percapita+LivingDonors+AllDonors+WaitinglistAdditions, data=mydata2)   #or VTABLE
#rate of living organ donations
#table(mydata2$rate_LivingDonors)

#The mean for all variables

# Some descriptive figures
View(mydata2)

linegraph2<-ggplot(mydata2, aes(x = Year, 
                y = LivingDonors, 
                group = State)) +
  geom_line() + 
  gghighlight(State=="Louisiana",
              label_params = list(fill = NA, alpha=1)) 
linegraph2+
  geom_line(aes(xintercept=2015),
            linetype=4, colour="black")+
  theme_classic()
```

#B


#C


```{r drop some states}

#drop those with zeros for living donors

`%!in%` <- Negate(`%in%`)
mydata2 <- mydata2 %>% filter(State  %!in% c("Alaska", "Idaho", "Montana",
                                             "Vermont", "Wyoming"))

```





#D


```{r Synthetic cont, message=FALSE, warning=FALSE, comment=NA}



mydata2$centered_year<-mydata2$Year-2015
# Construct Synthetic Controls
donation_out <-
  mydata2  %>%
  # initial the synthetic control object
  synthetic_control(outcome = LivingDonors, # outcome
                    unit = State, # unit index in the panel data
                    time = Year, # time index in the panel data
                    i_unit = "Louisiana", # unit where the intervention occurred
                    i_time = 2015, # time period when the intervention occurred
                    generate_placebos=T # generate placebo synthetic controls (for inference)
  ) %>%
    generate_predictor(time_window = 2000:2014,
    lagged_state_population=mean(Population, na.rm=T),
    lagged_state_GDP=mean(GDP, na.rm=T),
    lagged_totaldonations = mean(AllDonors, na.rm=T)) %>%
    # Generate the fitted weights for the synthetic control
  generate_weights(optimization_window = 2000:2014, # time to use in the optimization task
                   margin_ipop = .02,sigf_ipop = 7,bound_ipop = 6 # optimizer options
  ) %>%
  # Generate the synthetic control
  generate_control()

  
#relative weights

# Which states are we using, and what weights are they given? 
donation_out %>%
  grab_unit_weights() %>%
  mutate(weights = round(weight, digits = 4)) %>%
  select(unit, weights) %>%
  filter(weights>0.0001) %>%
  as.data.frame() %>%
  stargazer(summary = FALSE, rownames = FALSE,type="text")

# What about the independent variables?
donation_out %>% 
  plot_weights() + 
  labs(title="Synthetic Control Weights") 
```


#E

```{r Balance table,echo=TRUE, message=FALSE, comment=NA}

# Balance Table
donation_out %>%
  grab_balance_table() %>%
  mutate(difference = Louisiana - synthetic_Louisiana) %>%
  select(variable, Louisiana, synthetic_Louisiana, difference, donor_sample) %>%
  as.data.frame() %>%
  stargazer(summary = FALSE, rownames = FALSE, 
            caption = "Balance Table", 
            label = "balancetable",type="text") # 


donation_out %>% plot_trends() +
  scale_x_continuous(breaks = c(-15,-10,-5,0,5)) +
  labs(
    title = "Louisiana and Synthetic Louisiana",
    caption = "Timing of The donations",
    x="Years Relative to donations",
    y="rate of donations"
  ) 

```

#F Checking effectiveness
### Interpreting the Synthetic Control

So what's the effect? To find out, let's compare Louisiana and Synthetic Louisiana over time: 

```{r echo=TRUE, message=FALSE, comment=NA}


donation_out %>% plot_trends() +
  scale_x_continuous(breaks = c(-15,-10,-5,0,5)) +
  labs(
    title = "Louisiana and Synthetic Louisiana",
    caption = "Timing of The donations",
    x="Years Relative to donations",
    y="rate of donations"
  )

#So what's the effect? To find out, let's compare Louisiana and Synthetic Louisiana over time: 
# Plot Model Differences
donation_out %>% plot_differences() +
  scale_x_continuous(breaks = c(-15,-10,-5,0,5)) +
  labs(
   title = "Louisiana and Synthetic Louisiana",
    caption = "Timing of The donations",
    x="Years Relative to donations",
    y="rate of donations"
  ) 


```


#G


```{r Placebo tests, echo=TRUE, message=FALSE, comment=NA}
#i:
# Plot placebos of different states' assignments
### Inference
#How do we get a sense of whether the effect was significant? To do this, let's plot the placebo differences between each state and it's "synthetic state" in the world where the vaccine lottery never took place (hint: the lottery *never* took place in *any* of these states).


donation_out %>% plot_placebos() +
  scale_x_continuous(breaks = c(-15,-10,-5,0,5)) +
  labs(
    title = "Difference between State and Synthetic State: All States",
    caption = "Timing of The Donation",
    x="Years Relative to time of donations",
    y="Difference in rate donated"
  )


#ii

# This test shifts the pre-treatment window back five years


placebo_out <-
  mydata2 %>% 
  filter(Year <= 2018) %>% 
  # initial the synthetic control object
  synthetic_control(outcome = LivingDonors, # outcome
                    unit = State, # unit index in the panel data
                    time = Year, # time index in the panel data
                    i_unit = "Louisiana", # unit where the intervention occurred
                    i_time = 2018, # time period when the intervention occurred
                    generate_placebos=T # generate placebo synthetic controls (for inference)
  ) %>%
    generate_predictor(time_window = 2000:2017,
    lagged_state_population2=mean(Population, na.rm=T),
    lagged_state_GDP2=mean(GDP, na.rm=T),
    lagged_totaldonations2 = mean(AllDonors, na.rm=T)) %>%
    # Generate the fitted weights for the synthetic control
  generate_weights(optimization_window = 2000:2017, # time to use in the optimization task
                   margin_ipop = .02,sigf_ipop = 7,bound_ipop = 6 # optimizer options
  ) %>%
  # Generate the synthetic control
  generate_control()
  
  ###########################

#relative weights

# Which states are we using, and what weights are they given? 
placebo_out %>%
  grab_unit_weights() %>%
  mutate(weights = round(weight, digits = 4)) %>%
  select(unit, weights) %>%
  filter(weights>0.0001) %>%
  as.data.frame() %>%
  stargazer(summary = FALSE, rownames = FALSE,type="text")

# What about the independent variables?
placebo_out %>% 
  plot_weights() + 
  labs(title="Synthetic Control Weights") 




# Balance Table
placebo_out %>%
  grab_balance_table() %>%
  mutate(difference = Louisiana - synthetic_Louisiana) %>%
  select(variable, Louisiana, synthetic_Louisiana, difference, donor_sample) %>%
  as.data.frame() %>%
  stargazer(summary = FALSE, rownames = FALSE, 
            caption = "Balance Table", 
            label = "balancetable",type="text") # 


placebo_out %>% plot_trends() +
  scale_x_continuous(breaks = c(-15,-10,-5,0,5)) +
  labs(
    title = "Louisiana and Synthetic Louisiana",
    caption = "Timing of The donations",
    x="Years Relative to donations",
    y="rate of donations"
  ) 


#So what's the effect? To find out, let's compare Louisiana and Synthetic Louisiana over time: 
# Plot Model Differences
placebo_out %>% plot_differences() +
  scale_x_continuous(breaks = c(-15,-10,-5,0,5)) +
  labs(
   title = "Louisiana and Synthetic Louisiana",
    caption = "Timing of The donations",
    x="Years Relative to donations",
    y="rate of donations"
  ) 

#####################




  
  
  #####################

placebo_out %>% plot_trends()  + 
  labs(
    title = "Placebo Analysis: Louisiana and Synthetic Louisiana",
  caption = "Timing of The donations",
    x="Years Relative to donations",
    y="rate of donations"
  )


placebo_out %>% plot_differences()  + 
  labs(
    title = "Placebo Analysis:  Difference between Ohio and Synthetic Ohio",
   caption = "Timing of The donations",
    x="Years Relative to donations",
    y="rate of donations"
  )





```


#H




```{r References,  echo=TRUE, message=FALSE, comment=NA, results='asis'}
packages <- knitr::write_bib(file = 'packages.bib')
packages

```